### 2021-04-09 submission/manuscript.Rmd
Fix titlepage superscripts
Additions to the introduction

Sarah Tomkovich^1, Ana Taylor, Jacob King, Joanna Colovas, Lucas Bishop, Kathryn McBride, Sonya Royzenblat, Nicholas A. Lesniak, Ingrid L. Bergin^2, Patrick D. Schloss^1$\dagger$^$

@@

Antibiotics are a major risk factor for *Clostridioides difficile* infections (CDIs) because of their impact on the microbiota. However, antibiotics are not the only types of medications that disrupt the microbiota. Although, other medications such as proton pump inhibitors have been implicated as risk factors for CDIs through epidemiological studies, whether the association is due to their impact on the microbiome is still unclear [@Tomkovich2019]. Many of the non-antibiotic medications associated with CDIs are known to modulate intestinal motility, which in turn also strongly impacts microbiota composition and function.  
Interestingly our group has previously demonstrated that there are similar microbiota features in stool samples from CDI patients and patients with diarrhea that tested negative for *C. difficile*  with both groups having lower alpha diversity than non-diarrheal control samples [@Schubert2015]. These results led to our hypothesis that bacterial communities from patients experiencing diarrheal controls are susceptible, but have not been exposed to *C. difficile* spores.
To examine if diarrhea alone is sufficient to disrupt microbiota colonization resistance, we turned to osmotic laxatives, which increase intestinal motility and can lead to diarrhea. The ubiquitous osmotic laxative, polyethylene glycol (PEG) 3350 is found in Miralax, Nulytely, and Golytely and is also commonly used as bowel preparation for colonoscopies. Interestingly, previous studies has shown that treating mice with PEG alone rendered the mice susceptible to *C. difficile *infection, altered microbiota composition, reduced acetate and butyrate and altered the mucus barrier [@Kashyap2013; @Ferreyra2014; @Tropini2018; @VanInsberghe2020]. The mucus barrier is thought to mediate protection from *C. difficile* infections by protecting intestinal epithelial cells from the toxins produced by *C. difficile* (Ref).
PEG is also relevant in the context of treating recurrent CDIs via fecal microbiota transplant (FMT). For FMTs that are delivered via colonoscopy, patients typically undergo bowel preparation by taking an osmotic laxative prior to the procedure. Many of the FMT studies to date rationalize the use of laxatives (Ref) based on a case study with 2 pediatric patients where the authors suggested the laxative may help flush *C. difficile* spores and toxins from the intestine [@Liacouras1996].

* 2 main questions:
  + Does PEG 3350 prolong colonization and result in more severe C. difficile infections?
  + How does PEG 3350 influence C. difficile clearance and fecal microbiota transplant (FMT) in mice?

>>>>

Sarah Tomkovich^1^, Ana Taylor^1^, Jacob King^1^, Joanna Colovas^1^, Lucas Bishop^1^, Kathryn McBride^1^, Sonya Royzenblat^1^, Nicholas A. Lesniak^1^, Ingrid L. Bergin^2^, Patrick D. Schloss^1$\dagger$^

@@

Antibiotics are a major risk factor for *Clostridioides difficile* infections (CDIs) because of their impact on the microbiota [@Britton2014]. However, antibiotics are not the only types of medications that disrupt the microbiota [@Maier2018; @LeBastard2017; @VichVila2020]. Although, other medications have been implicated as risk factors for CDIs through epidemiological studies, whether the association is due to their impact on the microbiome is still unclear [@Oh2018; @Mora2012; @Nehra2018; @Krishna2013; @Tomkovich2019]. Many of the non-antibiotic medications associated with CDIs are known to modulate intestinal motility, which in turn also strongly impacts microbiota composition and function [@Vandeputte2015; @VujkovicCvijin2020].

Stool consistency often serves as an approximation of intestinal motility. Our group has shown that when *C. difficile* negative controls are separated into two groups based on stool consistency, there are microbiota features such as alpha diversity that overlap between samples from CDI patients and control patients with diarrhea [@Schubert2015]. These results led to our hypothesis that bacterial communities from patients experiencing diarrhea are susceptible, but may not have been exposed to *C. difficile* spores.

Osmotic laxatives can lead to diarrhea depending on the administered dose and temporarily disrupts the intestinal microbiota in humans [@Nagata2019]. The ubiquitous osmotic laxative, polyethylene glycol (PEG) 3350 is found in Miralax, Nulytely, and Golytely and is also commonly used as bowel preparation for colonoscopies. Interestingly, previous studies have shown that treating mice with PEG alone rendered the mice susceptible to *C. difficile* infection, altered microbiota composition, reduced acetate and butyrate and altered the mucus barrier [@Kashyap2013; @Ferreyra2014; @Tropini2018; @VanInsberghe2020]. The mucus barrier is thought to mediate protection from *C. difficile* infections by protecting intestinal epithelial cells from the toxins produced by *C. difficile* (Ref). However, whether laxative results in more severe CDIs in mice is unclear.

Beyond susceptibility, PEG is also relevant in the context of treating recurrent CDIs via fecal microbiota transplant (FMT) where a healthy microbiota is administered to the patient to restore colonization. For FMTs that are delivered via colonoscopy, patients typically undergo bowel preparation by taking an osmotic laxative prior to the procedure. Many of the FMT studies to date rationalize the use of laxatives (Ref) based on a 1996 case study with 2 pediatric patients where the authors suggested the laxative may help flush *C. difficile* spores and toxins from the intestine [@Liacouras1996].

In the past, our group has used C57BL6 mice to characterize how antibiotics disrupt the microbiota and influence *C. difficile* susceptibility and clearance [ref]. Although, two groups have now shown PEG treatment alone renders mice susceptible to *C. difficile*, these studies have raised additional questions that should be addressed given their relevance to CDIs. Here, we used our C57BL/6 clindamycin model as a control group to characterize how long PEG-treated mice remain susceptible, whether PEG treatment results in sustained *C. difficile* colonization, and if PEG treatment post-CDI can promote *C. difficile* clearance.

### 2021-04-09 submission/manuscript.tex
Additions to the introduction

Antibiotics are a major risk factor for \emph{Clostridioides difficile}
infections (CDIs) because of their impact on the microbiota. However,
antibiotics are not the only types of medications that disrupt the
microbiota. Although, other medications such as proton pump inhibitors
have been implicated as risk factors for CDIs through epidemiological
studies, whether the association is due to their impact on the
microbiome is still unclear (1). Many of the non-antibiotic medications
associated with CDIs are known to modulate intestinal motility, which in
turn also strongly impacts microbiota composition and function.\\
Interestingly our group has previously demonstrated that there are
similar microbiota features in stool samples from CDI patients and
patients with diarrhea that tested negative for \emph{C. difficile} with
both groups having lower alpha diversity than non-diarrheal control
samples (2). These results led to our hypothesis that bacterial
communities from patients experiencing diarrheal controls are
susceptible, but have not been exposed to \emph{C. difficile} spores. To
examine if diarrhea alone is sufficient to disrupt microbiota
colonization resistance, we turned to osmotic laxatives, which increase
intestinal motility and can lead to diarrhea. The ubiquitous osmotic
laxative, polyethylene glycol (PEG) 3350 is found in Miralax, Nulytely,
and Golytely and is also commonly used as bowel preparation for
colonoscopies. Interestingly, previous studies has shown that treating
mice with PEG alone rendered the mice susceptible to \emph{C. difficile
}infection, altered microbiota composition, reduced acetate and butyrate
and altered the mucus barrier (3--6). The mucus barrier is thought to
mediate protection from \emph{C. difficile} infections by protecting
intestinal epithelial cells from the toxins produced by \emph{C.
difficile} (Ref). PEG is also relevant in the context of treating
recurrent CDIs via fecal microbiota transplant (FMT). For FMTs that are
delivered via colonoscopy, patients typically undergo bowel preparation
by taking an osmotic laxative prior to the procedure. Many of the FMT
studies to date rationalize the use of laxatives (Ref) based on a case
study with 2 pediatric patients where the authors suggested the laxative
may help flush \emph{C. difficile} spores and toxins from the intestine
(7).

\begin{itemize}
\tightlist
\item
  2 main questions:

  \begin{itemize}
  \tightlist
  \item
    Does PEG 3350 prolong colonization and result in more severe C.
    difficile infections?
  \item
    How does PEG 3350 influence C. difficile clearance and fecal
    microbiota transplant (FMT) in mice?
  \end{itemize}
\end{itemize}

@@

\textbf{Laxative treatment alone leads to prolonged \emph{C. difficile}
colonization in mice.} We compared PEG-treated mice to our standard 10
mg/kg clindamycin treatment, which temporarily renders the mice
susceptible to \emph{C. difficile}, with mice typically clearing C.
difficile within 10 days post-infection (1, 8). All PEG-treated mice
were administered a 15\% PEG solution in the drinking water for 5-days,
one group was also treated with clindamycin, and one group was allowed
to recover for 10 days prior to challenge (Fig. 1A). After PEG and/or
antibiotic treatment all mice were challenged with 10\textsuperscript{3}
\emph{C. difficile} 630 spores.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Tomkovich2019}{}%
1. \textbf{Tomkovich S}, \textbf{Lesniak NA}, \textbf{Li Y},
\textbf{Bishop L}, \textbf{Fitzgerald MJ}, \textbf{Schloss PD}. 2019.
The proton pump inhibitor omeprazole does not promote
\emph{Clostridioides difficile} colonization in a murine model. mSphere
\textbf{4}.
doi:\href{https://doi.org/10.1128/msphere.00693-19}{10.1128/msphere.00693-19}.

\leavevmode\hypertarget{ref-Schubert2015}{}%
2. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against \emph{Clostridium
difficile}. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\leavevmode\hypertarget{ref-Kashyap2013}{}%
3. \textbf{Kashyap PC}, \textbf{Marcobal A}, \textbf{Ursell LK},
\textbf{Larauche M}, \textbf{Duboc H}, \textbf{Earle KA},
\textbf{Sonnenburg ED}, \textbf{Ferreyra JA}, \textbf{Higginbottom SK},
\textbf{Million M}, \textbf{Tache Y}, \textbf{Pasricha PJ},
\textbf{Knight R}, \textbf{Farrugia G}, \textbf{Sonnenburg JL}. 2013.
Complex interactions among diet, gastrointestinal transit, and gut
microbiota in humanized mice. Gastroenterology \textbf{144}:967--977.
doi:\href{https://doi.org/10.1053/j.gastro.2013.01.047}{10.1053/j.gastro.2013.01.047}.

\leavevmode\hypertarget{ref-Ferreyra2014}{}%
4. \textbf{Ferreyra JA}, \textbf{Wu KJ}, \textbf{Hryckowian AJ},
\textbf{Bouley DM}, \textbf{Weimer BC}, \textbf{Sonnenburg JL}. 2014.
Gut microbiota-produced succinate promotes c.~difficile infection after
antibiotic treatment or motility disturbance. Cell Host \& Microbe
\textbf{16}:770--777.
doi:\href{https://doi.org/10.1016/j.chom.2014.11.003}{10.1016/j.chom.2014.11.003}.

\leavevmode\hypertarget{ref-Tropini2018}{}%
5. \textbf{Tropini C}, \textbf{Moss EL}, \textbf{Merrill BD}, \textbf{Ng
KM}, \textbf{Higginbottom SK}, \textbf{Casavant EP}, \textbf{Gonzalez
CG}, \textbf{Fremin B}, \textbf{Bouley DM}, \textbf{Elias JE},
\textbf{Bhatt AS}, \textbf{Huang KC}, \textbf{Sonnenburg JL}. 2018.
Transient osmotic perturbation causes long-term alteration to the gut
microbiota. Cell \textbf{173}:1742--1754.e17.
doi:\href{https://doi.org/10.1016/j.cell.2018.05.008}{10.1016/j.cell.2018.05.008}.

\leavevmode\hypertarget{ref-VanInsberghe2020}{}%
6. \textbf{VanInsberghe D}, \textbf{Elsherbini JA}, \textbf{Varian B},
\textbf{Poutahidis T}, \textbf{Erdman S}, \textbf{Polz MF}. 2020.
Diarrhoeal events can trigger long-term clostridium difficile
colonization with recurrent blooms. Nature Microbiology
\textbf{5}:642--650.
doi:\href{https://doi.org/10.1038/s41564-020-0668-2}{10.1038/s41564-020-0668-2}.

\leavevmode\hypertarget{ref-Liacouras1996}{}%
7. \textbf{Liacouras CA}, \textbf{Piccoli DA}. 1996. Whole-bowel
irrigation as an adjunct to the treatment of chronic, relapsing
clostridium difficile colitis. Journal of Clinical Gastroenterology
\textbf{22}:186--189.
doi:\href{https://doi.org/10.1097/00004836-199604000-00007}{10.1097/00004836-199604000-00007}.

\leavevmode\hypertarget{ref-Tomkovich2020}{}%
8. \textbf{Tomkovich S}, \textbf{Stough JMA}, \textbf{Bishop L},
\textbf{Schloss PD}. 2020. The initial gut microbiota and response to
antibiotic perturbation influence clostridioides difficile clearance in
mice. mSphere \textbf{5}.
doi:\href{https://doi.org/10.1128/msphere.00869-20}{10.1128/msphere.00869-20}.

>>>>

Antibiotics are a major risk factor for \emph{Clostridioides difficile}
infections (CDIs) because of their impact on the microbiota (1).
However, antibiotics are not the only types of medications that disrupt
the microbiota (2--4). Although, other medications have been implicated
as risk factors for CDIs through epidemiological studies, whether the
association is due to their impact on the microbiome is still unclear
(5--9). Many of the non-antibiotic medications associated with CDIs are
known to modulate intestinal motility, which in turn also strongly
impacts microbiota composition and function (10, 11).

Stool consistency often serves as an approximation of intestinal
motility. Our group has shown that when \emph{C. difficile} negative
controls are separated into two groups based on stool consistency, there
are microbiota features such as alpha diversity that overlap between
samples from CDI patients and control patients with diarrhea (12). These
results led to our hypothesis that bacterial communities from patients
experiencing diarrhea are susceptible, but may not have been exposed to
\emph{C. difficile} spores.

Osmotic laxatives can lead to diarrhea depending on the administered
dose and temporarily disrupts the intestinal microbiota in humans (13).
The ubiquitous osmotic laxative, polyethylene glycol (PEG) 3350 is found
in Miralax, Nulytely, and Golytely and is also commonly used as bowel
preparation for colonoscopies. Interestingly, previous studies have
shown that treating mice with PEG alone rendered the mice susceptible to
\emph{C. difficile} infection, altered microbiota composition, reduced
acetate and butyrate and altered the mucus barrier (14--17). The mucus
barrier is thought to mediate protection from \emph{C. difficile}
infections by protecting intestinal epithelial cells from the toxins
produced by \emph{C. difficile} (Ref). However, whether laxative results
in more severe CDIs in mice is unclear.

Beyond susceptibility, PEG is also relevant in the context of treating
recurrent CDIs via fecal microbiota transplant (FMT) where a healthy
microbiota is administered to the patient to restore colonization. For
FMTs that are delivered via colonoscopy, patients typically undergo
bowel preparation by taking an osmotic laxative prior to the procedure.
Many of the FMT studies to date rationalize the use of laxatives (Ref)
based on a 1996 case study with 2 pediatric patients where the authors
suggested the laxative may help flush \emph{C. difficile} spores and
toxins from the intestine (18).

In the past, our group has used C57BL6 mice to characterize how
antibiotics disrupt the microbiota and influence \emph{C. difficile}
susceptibility and clearance {[}ref{]}. Although, two groups have now
shown PEG treatment alone renders mice susceptible to \emph{C.
difficile}, these studies have raised additional questions that should
be addressed given their relevance to CDIs. Here, we used our C57BL/6
clindamycin model as a control group to characterize how long
PEG-treated mice remain susceptible, whether PEG treatment results in
sustained \emph{C. difficile} colonization, and if PEG treatment
post-CDI can promote \emph{C. difficile} clearance.

@@

\textbf{Laxative treatment alone leads to prolonged \emph{C. difficile}
colonization in mice.} We compared PEG-treated mice to our standard 10
mg/kg clindamycin treatment, which temporarily renders the mice
susceptible to \emph{C. difficile}, with mice typically clearing C.
difficile within 10 days post-infection (9, 19). All PEG-treated mice
were administered a 15\% PEG solution in the drinking water for 5-days,
one group was also treated with clindamycin, and one group was allowed
to recover for 10 days prior to challenge (Fig. 1A). After PEG and/or
antibiotic treatment all mice were challenged with 10\textsuperscript{3}
\emph{C. difficile} 630 spores.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Britton2014}{}%
1. \textbf{Britton RA}, \textbf{Young VB}. 2014. Role of the intestinal
microbiota in resistance to colonization by clostridium difficile.
Gastroenterology \textbf{146}:1547--1553.
doi:\href{https://doi.org/10.1053/j.gastro.2014.01.059}{10.1053/j.gastro.2014.01.059}.

\leavevmode\hypertarget{ref-Maier2018}{}%
2. \textbf{Maier L}, \textbf{Pruteanu M}, \textbf{Kuhn M},
\textbf{Zeller G}, \textbf{Telzerow A}, \textbf{Anderson EE},
\textbf{Brochado AR}, \textbf{Fernandez KC}, \textbf{Dose H},
\textbf{Mori H}, \textbf{Patil KR}, \textbf{Bork P}, \textbf{Typas A}.
2018. Extensive impact of non-antibiotic drugs on human gut bacteria.
Nature \textbf{555}:623--628.
doi:\href{https://doi.org/10.1038/nature25979}{10.1038/nature25979}.

\leavevmode\hypertarget{ref-LeBastard2017}{}%
3. \textbf{Bastard QL}, \textbf{Al-Ghalith GA}, \textbf{Grégoire M},
\textbf{Chapelet G}, \textbf{Javaudin F}, \textbf{Dailly E},
\textbf{Batard E}, \textbf{Knights D}, \textbf{Montassier E}. 2017.
Systematic review: Human gut dysbiosis induced by non-antibiotic
prescription medications. Alimentary Pharmacology \& Therapeutics
\textbf{47}:332--345.
doi:\href{https://doi.org/10.1111/apt.14451}{10.1111/apt.14451}.

\leavevmode\hypertarget{ref-VichVila2020}{}%
4. \textbf{Vila AV}, \textbf{Collij V}, \textbf{Sanna S}, \textbf{Sinha
T}, \textbf{Imhann F}, \textbf{Bourgonje AR}, \textbf{Mujagic Z},
\textbf{Jonkers DMAE}, \textbf{Masclee AAM}, \textbf{Fu J},
\textbf{Kurilshikov A}, \textbf{Wijmenga C}, \textbf{Zhernakova A},
\textbf{Weersma RK}. 2020. Impact of commonly used drugs on the
composition and metabolic function of the gut microbiota. Nature
Communications \textbf{11}.
doi:\href{https://doi.org/10.1038/s41467-019-14177-z}{10.1038/s41467-019-14177-z}.

\leavevmode\hypertarget{ref-Oh2018}{}%
5. \textbf{Oh J}, \textbf{Makar M}, \textbf{Fusco C}, \textbf{McCaffrey
R}, \textbf{Rao K}, \textbf{Ryan EE}, \textbf{Washer L}, \textbf{West
LR}, \textbf{Young VB}, \textbf{Guttag J}, \textbf{Hooper DC},
\textbf{Shenoy ES}, \textbf{Wiens J}. 2018. A generalizable, data-driven
approach to predict daily risk ofClostridium difficileInfection at two
large academic health centers. Infection Control \& Hospital
Epidemiology \textbf{39}:425--433.
doi:\href{https://doi.org/10.1017/ice.2018.16}{10.1017/ice.2018.16}.

\leavevmode\hypertarget{ref-Mora2012}{}%
6. \textbf{Mora AL}, \textbf{Salazar M}, \textbf{Pablo-Caeiro J},
\textbf{Frost CP}, \textbf{Yadav Y}, \textbf{DuPont HL}, \textbf{Garey
KW}. 2012. Moderate to high use of opioid analgesics are associated with
an increased risk of clostridium difficile infection. The American
Journal of the Medical Sciences \textbf{343}:277--280.
doi:\href{https://doi.org/10.1097/maj.0b013e31822f42eb}{10.1097/maj.0b013e31822f42eb}.

\leavevmode\hypertarget{ref-Nehra2018}{}%
7. \textbf{Nehra AK}, \textbf{Alexander JA}, \textbf{Loftus CG},
\textbf{Nehra V}. 2018. Proton pump inhibitors: Review of emerging
concerns. Mayo Clinic Proceedings \textbf{93}:240--246.
doi:\href{https://doi.org/10.1016/j.mayocp.2017.10.022}{10.1016/j.mayocp.2017.10.022}.

\leavevmode\hypertarget{ref-Krishna2013}{}%
8. \textbf{Krishna SG}, \textbf{Zhao W}, \textbf{Apewokin SK},
\textbf{Krishna K}, \textbf{Chepyala P}, \textbf{Anaissie EJ}. 2013.
Risk factors, preemptive therapy, and antiperistaltic agents
forClostridium difficileinfection in cancer patients. Transplant
Infectious Disease n/a--n/a.
doi:\href{https://doi.org/10.1111/tid.12112}{10.1111/tid.12112}.

\leavevmode\hypertarget{ref-Tomkovich2019}{}%
9. \textbf{Tomkovich S}, \textbf{Lesniak NA}, \textbf{Li Y},
\textbf{Bishop L}, \textbf{Fitzgerald MJ}, \textbf{Schloss PD}. 2019.
The proton pump inhibitor omeprazole does not promote
\emph{Clostridioides difficile} colonization in a murine model. mSphere
\textbf{4}.
doi:\href{https://doi.org/10.1128/msphere.00693-19}{10.1128/msphere.00693-19}.

\leavevmode\hypertarget{ref-Vandeputte2015}{}%
10. \textbf{Vandeputte D}, \textbf{Falony G}, \textbf{Vieira-Silva S},
\textbf{Tito RY}, \textbf{Joossens M}, \textbf{Raes J}. 2015. Stool
consistency is strongly associated with gut microbiota richness and
composition, enterotypes and bacterial growth rates. Gut
\textbf{65}:57--62.
doi:\href{https://doi.org/10.1136/gutjnl-2015-309618}{10.1136/gutjnl-2015-309618}.

\leavevmode\hypertarget{ref-VujkovicCvijin2020}{}%
11. \textbf{Vujkovic-Cvijin I}, \textbf{Sklar J}, \textbf{Jiang L},
\textbf{Natarajan L}, \textbf{Knight R}, \textbf{Belkaid Y}. 2020. Host
variables confound gut microbiota studies of human disease. Nature
\textbf{587}:448--454.
doi:\href{https://doi.org/10.1038/s41586-020-2881-9}{10.1038/s41586-020-2881-9}.

\leavevmode\hypertarget{ref-Schubert2015}{}%
12. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against \emph{Clostridium
difficile}. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\leavevmode\hypertarget{ref-Nagata2019}{}%
13. \textbf{Nagata N}, \textbf{Tohya M}, \textbf{Fukuda S}, \textbf{Suda
W}, \textbf{Nishijima S}, \textbf{Takeuchi F}, \textbf{Ohsugi M},
\textbf{Tsujimoto T}, \textbf{Nakamura T}, \textbf{Shimomura A},
\textbf{Yanagisawa N}, \textbf{Hisada Y}, \textbf{Watanabe K},
\textbf{Imbe K}, \textbf{Akiyama J}, \textbf{Mizokami M},
\textbf{Miyoshi-Akiyama T}, \textbf{Uemura N}, \textbf{Hattori M}. 2019.
Effects of bowel preparation on the human gut microbiome and metabolome.
Scientific Reports \textbf{9}.
doi:\href{https://doi.org/10.1038/s41598-019-40182-9}{10.1038/s41598-019-40182-9}.

\leavevmode\hypertarget{ref-Kashyap2013}{}%
14. \textbf{Kashyap PC}, \textbf{Marcobal A}, \textbf{Ursell LK},
\textbf{Larauche M}, \textbf{Duboc H}, \textbf{Earle KA},
\textbf{Sonnenburg ED}, \textbf{Ferreyra JA}, \textbf{Higginbottom SK},
\textbf{Million M}, \textbf{Tache Y}, \textbf{Pasricha PJ},
\textbf{Knight R}, \textbf{Farrugia G}, \textbf{Sonnenburg JL}. 2013.
Complex interactions among diet, gastrointestinal transit, and gut
microbiota in humanized mice. Gastroenterology \textbf{144}:967--977.
doi:\href{https://doi.org/10.1053/j.gastro.2013.01.047}{10.1053/j.gastro.2013.01.047}.

\leavevmode\hypertarget{ref-Ferreyra2014}{}%
15. \textbf{Ferreyra JA}, \textbf{Wu KJ}, \textbf{Hryckowian AJ},
\textbf{Bouley DM}, \textbf{Weimer BC}, \textbf{Sonnenburg JL}. 2014.
Gut microbiota-produced succinate promotes c.~difficile infection after
antibiotic treatment or motility disturbance. Cell Host \& Microbe
\textbf{16}:770--777.
doi:\href{https://doi.org/10.1016/j.chom.2014.11.003}{10.1016/j.chom.2014.11.003}.

\leavevmode\hypertarget{ref-Tropini2018}{}%
16. \textbf{Tropini C}, \textbf{Moss EL}, \textbf{Merrill BD},
\textbf{Ng KM}, \textbf{Higginbottom SK}, \textbf{Casavant EP},
\textbf{Gonzalez CG}, \textbf{Fremin B}, \textbf{Bouley DM},
\textbf{Elias JE}, \textbf{Bhatt AS}, \textbf{Huang KC},
\textbf{Sonnenburg JL}. 2018. Transient osmotic perturbation causes
long-term alteration to the gut microbiota. Cell
\textbf{173}:1742--1754.e17.
doi:\href{https://doi.org/10.1016/j.cell.2018.05.008}{10.1016/j.cell.2018.05.008}.

\leavevmode\hypertarget{ref-VanInsberghe2020}{}%
17. \textbf{VanInsberghe D}, \textbf{Elsherbini JA}, \textbf{Varian B},
\textbf{Poutahidis T}, \textbf{Erdman S}, \textbf{Polz MF}. 2020.
Diarrhoeal events can trigger long-term clostridium difficile
colonization with recurrent blooms. Nature Microbiology
\textbf{5}:642--650.
doi:\href{https://doi.org/10.1038/s41564-020-0668-2}{10.1038/s41564-020-0668-2}.

\leavevmode\hypertarget{ref-Liacouras1996}{}%
18. \textbf{Liacouras CA}, \textbf{Piccoli DA}. 1996. Whole-bowel
irrigation as an adjunct to the treatment of chronic, relapsing
clostridium difficile colitis. Journal of Clinical Gastroenterology
\textbf{22}:186--189.
doi:\href{https://doi.org/10.1097/00004836-199604000-00007}{10.1097/00004836-199604000-00007}.

\leavevmode\hypertarget{ref-Tomkovich2020}{}%
19. \textbf{Tomkovich S}, \textbf{Stough JMA}, \textbf{Bishop L},
\textbf{Schloss PD}. 2020. The initial gut microbiota and response to
antibiotic perturbation influence clostridioides difficile clearance in
mice. mSphere \textbf{5}.
doi:\href{https://doi.org/10.1128/msphere.00869-20}{10.1128/msphere.00869-20}.
